Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-10-08 SupportingInformation.Rmd
modified

<!--Display model selection results for BMR:birds data in a table-->
```{r tab_bmr_mammals_fit, echo=F, results='markup'}
h<-BivMS_mammals$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='pandoc',
             caption = "Fitting info for **mammals' BMR dataset** . \\label{tab_bmr_birds_fit}",
             booktabs=T)
```

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf}
\caption[Testing GOF between noise-copula parameters and estimated population copula parameters with $2500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{2500} points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF_2500}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Cause4copula_results/SClayton_kendall.pdf}
\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e. Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_SC_kend}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Cause4copula_results/Clayton_spearman.pdf}
\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Spearman's Rho}]{Results of comparison between 3 types of correlation (i.e. Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Spearman's Rho} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_C_spear}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Cause4copula_results/SClayton_spearman.pdf}
\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Spearman's Rho}]{Results of comparison between 3 types of correlation (i.e. Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Spearman's Rho} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_SC_spear}}
\end{center}
\end{figure}

@@

```{r tab_temp_aphid_bestcop, echo=F, results="markup"}
knitr::kable(RES_temp_aphid$info_ord_copname[,,1], #longtable = T,
             format='pandoc',
             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **negative average winter temperature data (1958-2013)**. Copulas' abbreviations were given in \\nameref{MT-Methods}. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.  \\label{tab_temp_aphid_bestcop}",
             booktabs=TRUE
             )
```

@@

\newpage
```{r tab_temp_aphid_LTmUT, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw,
             digits = 5,
             format='pandoc',
             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTmUT}",
             booktabs=TRUE
             )
```

<!--# Cause of non-Normal copula structure : the Moran effect and copula structure in populations \label{Cause_Moran}-->

@@

```{r tab_stat_npa_asym_4,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[4]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[4]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_4}"))
```

@@

#  Need to think about mechanisms for biogeochemical, evolutionary, and community data

@@

pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_kendall.pdf",height=20,width=20)
op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
#cat("Now for kendall preserving surrogs : pns","\n")
for(i in splist){
  #cat("sp=",i,"\n")
  loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=i,data_pt_thrs=45)
  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_plankton_north_sea),loclist=loclist,
                        numsurrog=numsurrog,
                        ploton=T,
                        corpres="kendall") 
  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
  skw_plankton_kendall$realskw[i]<-ans$realskw
  skw_plankton_kendall$p_left[i]<-p_left
  skw_plankton_kendall$p_right[i]<-p_right
  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
}
par(op)
dev.off()
saveRDS(skw_plankton_kendall,"./Results/skewness_results/skewness_plankton_north_sea/skw_plankton_kendall.RDS")
```

@@

#Now for sp=6
sp<-6
pdf("./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf",height=5,width=5)
ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=sp,data_pt_thrs = 30,ploton=T,tag="mncurv",labelon = F)
mtext(paste0("p= ",round(ans$p_left_mncurv,3)),side = 3, line=-2, adj=0.5, col="blue")
dev.off()

@@

# Consequence of non-Normal copula structure : Effect on Taylor's law \label{Consequence_TL}

Spatial Taylor's law, which shows linear relationship between `log`(variance) and `log`(mean) of population densities of a species from different locations, is a well-known pattern found in ecology and related fields [@Taylor1961, @Taylor1988, @Cohen2013, @Doring2015, @Kuo2016, @Xu2016, @Reuman2017]. Here, we examined the nature of the slope and how it changed if population time series (taken as $50$ years) from different locations (say, $numloc=25$) had some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula. First, we simulated (N $\times$ n $\times$ nr) dimensional normal copula with covariance $\rho =0.7$, where $(N=50,n=25,nr=10)$ are the length of timeseries, number of sampling locations and number of replica for that copula, respectively. Then we generated Kendall-correlation preserving \textbf{C}layton, \textbf{F}rank and $\textbf{S}$urvival \textbf{C}layton surrogates with same dimension of \textbf{N}ormal. With these four different types of copula we computed `log`(spatial variance) and `log`(mean) and plotted in Fig.\ref{fig_spTL}(A-D). We also tested statistically that whether linearity, homoskedasticity is maintained in each copula dataset or not and plotted our results alongwith root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I) (For code see [`copsurrognd.R`](https://github.com/sghosh89/BIVAN/blob/master/copsurrognd.R) and [`TaylorsLawTools.R`](https://github.com/sghosh89/BIVAN/blob/master/TaylorsLawTools.R)).

>>>>

<!--Display model selection results for BMR:birds data in a table-->
```{r tab_bmr_mammals_fit, echo=F, results='markup'}
h<-BivMS_mammals$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='pandoc',
             caption = "Fitting info for **mammals' BMR dataset** . \\label{tab_bmr_mammals_fit}",
             booktabs=T)
```

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf}
\caption[Testing GOF between noise-copula parameters and estimated population copula parameters with $2500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{2500} points for both noise and population data for comparison). p-values $> 0.05$ (above horizontal dashed line) indicates good fit between noise and population copulae. \label{fig_Cause4copula_params_GOF_2500}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Cause4copula_results/SClayton_kendall.pdf}
\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e. Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison). p-values $< 0.05$ (below horizontal dashed line) indicates significantly similar estimate between noise and population copulae. \label{fig_Cause4copula_SC_kend}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Cause4copula_results/Clayton_spearman.pdf}
\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Spearman's Rho}]{Results of comparison between 3 types of correlation (i.e. Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Spearman's Rho} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison). p-values $< 0.05$ (below horizontal dashed line) indicates significantly similar estimate between noise and population copulae. \label{fig_Cause4copula_C_spear}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Cause4copula_results/SClayton_spearman.pdf}
\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Spearman's Rho}]{Results of comparison between 3 types of correlation (i.e. Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Spearman's Rho} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison). p-values $< 0.05$ (below horizontal dashed line) indicates significantly similar estimate between noise and population copulae. \label{fig_Cause4copula_SC_spear}}
\end{center}
\end{figure}

@@

<!--
```{r tab_temp_aphid_bestcop, echo=F, results="markup"}
knitr::kable(RES_temp_aphid$info_ord_copname[,,1], #longtable = T,
             format='pandoc',
             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **negative average winter temperature data (1958-2013)**. Copulas' abbreviations were given in \\nameref{MT-Methods}. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.  \\label{tab_temp_aphid_bestcop}",
             booktabs=TRUE
             )
```

@@

\newpage
```{r tab_temp_aphid_LTmUT, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw,
             digits = 5,
             format='pandoc',
             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTmUT}",
             booktabs=TRUE
             )
```
-->
<!--# Cause of non-Normal copula structure : the Moran effect and copula structure in populations \label{Cause_Moran}-->

@@

\newpage
```{r tab_stat_npa_asym_4,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[4]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[4]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_4}"))
```

@@

@@

pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_kendall.pdf",height=24,width=24)
op<-par(mfrow=c(4,6),mar=c(4.2,5,1,1),mgp=c(3,1,0))
#cat("Now for kendall preserving surrogs : pns","\n")
for(i in splist){
  #cat("sp=",i,"\n")
  loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=i,data_pt_thrs=45)
  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_plankton_north_sea),loclist=loclist,
                        numsurrog=numsurrog,
                        ploton=T,
                        corpres="kendall") 
  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
  skw_plankton_kendall$realskw[i]<-ans$realskw
  skw_plankton_kendall$p_left[i]<-p_left
  skw_plankton_kendall$p_right[i]<-p_right
  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
}
par(op)
dev.off()
saveRDS(skw_plankton_kendall,"./Results/skewness_results/skewness_plankton_north_sea/skw_plankton_kendall.RDS")
```

@@

#Now for sp=6
sp<-6
pdf("./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf",height=5,width=5)
op<-par(mar=c(4,5,1,1),mgp=c(3,1,0))
ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=sp,data_pt_thrs = 30,ploton=T,tag="mncurv",labelon = F)
mtext(paste0("p= ",round(ans$p_left_mncurv,3)),side = 3, line=-2, adj=0.5, col="blue")
par(op)
dev.off()

@@

# Consequence of non-Normal copula structure : Effect on Taylor's law \label{Consequence_TL_simulation}

Here, we examined the nature of the slope and how it changed if population time series (taken as $50$ years) from different locations (say, $numloc=25$) had some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula. First, we simulated (N $\times$ n $\times$ nr) dimensional normal copula with covariance $\rho =0.7$, where $(N=50,n=25,nr=10)$ are the length of timeseries, number of sampling locations and number of replica for that copula, respectively. Then we generated Kendall-correlation preserving \textbf{C}layton, \textbf{F}rank and $\textbf{S}$urvival \textbf{C}layton surrogates with same dimension of \textbf{N}ormal. With these four different types of copula we computed `log`(spatial variance) and `log`(mean) and plotted in Fig.\ref{fig_spTL}(A-D). We also tested statistically that whether linearity, homoskedasticity is maintained in each copula dataset or not and plotted our results alongwith root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I). We did a quadratic regression with each simulated copula for all types and measured quadratic coefficient and mean curvature for the regression line which is plotted in Fig.\ref{fig_spTL}(J-K) (For code see [`copsurrognd.R`](https://github.com/sghosh89/BIVAN/blob/master/copsurrognd.R) and [`TaylorsLawTools.R`](https://github.com/sghosh89/BIVAN/blob/master/TaylorsLawTools.R)).
